Measuring the Type Ia supernova delay time distribution via host galaxy rates in the Dark Energy Survey

### 2021-01-18 draft.tex

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citetalias{V20} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, we assign each SN in the sample described in Section \ref{subsec:host_sample} a weight equal to the inverse of the spectroscopic efficiency for its host. This is equivalent to assuming that an efficiency of 0.1 means that for every ten SNe with hosts at that magnitude, only 1 will make it into the sample.

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8 \mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the MS and become a WD, such as 40~Myr \citep{Maoz2012,Graur2014,Graur2018}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Asier2020}. We perform two fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, and one with it as a free parameter. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.2)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature.

@@

The majority of works in the field have constrained the DTD power-law slope, and many also estimate its normalisation. Conversely, few have attempted to constrain the time after a burst of star-formation at which SNe Ia begin to explode -- the prompt time $t_{\mathrm{p}}$. In most cases, $t_{\mathrm{p}}$ has been fixed at some fiducial value such as 40~Myr, which is derived from the lifetime of 8~M$_{\odot}$ stars. \citet{Asier2020} included $t_{\mathrm{p}}$ (which they denote $\Delta$) in their fit, and find a value of $50_{-35}^{+100}$~Myr which is consistent with both 40~Myr as well as our best fitting value of $90 \pm 50$~Myr. \citet{Heringer2019} also present fits with varying prompt times, although they don't fit for the parameter itself. They find that varying $t_{\mathrm{p}}$ changes the recovered normalisation, which is consistent with the degeneracy seen in our results (Fig. \ref{fig:corner_beta_norm_tp}). We therefore report results that are consistent with the fiducial value of 40~Myr.

@@

Low mass galaxies are highly sensitive to environmental effects and may often be found in the so-called starburst state with abnormally high (above the SFMS) levels of star formation that could drive environmental evolution \citep[e.g.][]{Vader1986,Marlowe1995,Tremonti2004,Lee2006}, although studies have shown that only 6 percent of dwarf galaxies are starbursts \citep{Lee2009}. It is possible that the averaging effects of the mass assembly model overestimate the true, instantaneous SFR for a significant fraction of the dwarf galaxies. Since the SN Ia rate is dominated by short delay times in these populations, an overestimated SFR in the model would lead to an overestimate in the SN Ia rate.

@@

One possible cause of the discrepant SN Ia rate at high stellar mass is the effect of stellar metallicity. Metallicity has been previously invoked to explain irregularities or divergences from the fiducial DTD \citep[e.g.][]{Strolger2010,Meng2011,Kistler2013}. Metallicity may affect the observed rate of SNe in two ways: firstly, it can affect the time taken for a star to evolve along the main sequence ($t_{\mathrm{MS}}$) -- Stellar metallicity has varying effects on the MS lifetime of stars \citep[e.g.][]{Georgy2007,Amard2020}, which is also dependent upon initial rotation and degree of mixing; secondly it can affect the time taken from WD formation to SN Ia explosion -- low metallicity stars should produce higher-mass WDs \citep[e.g.][]{Umeda1999,Marigo2007}, resulting in a higher SN Ia rate \citep{Kistler2013}.

>>>>

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citetalias{V20} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are eCastrillo to obtain spectroscopic redshifts for, we assign each SN in the sample described in Section \ref{subsec:host_sample} a weight equal to the inverse of the spectroscopic efficiency for its host. This is equivalent to assuming that an efficiency of 0.1 means that for every ten SNe with hosts at that magnitude, only 1 will make it into the sample.

@@

SNe Ia do not occur instantaneously after an episode of star formation. Instead stars with zero-age main-sequence (ZAMS) masses less than $8 \mathrm{M}_{\odot}$ take time to evolve along the main sequence and form WDs and this time is not well constrained. In many works this ``prompt time" $t_{\mathrm{p}}$ is fixed to some value expected to be the minimum time for a star to evolve off the MS and become a WD, such as 40~Myr \citep{Maoz2012,Graur2014,Graur2018}. Other studies have left $t_{\mathrm{p}}$ as a free parameter \citep{Heringer2019,Castrillo2020}. We perform two fits: one with $t_{\mathrm{p}}$ fixed at 40 Myr, and one with it as a free parameter. We fit $t_{\mathrm{p}}$ in log space and adopt a Gaussian prior with hyper-parameters $p(\log(t_{\mathrm{p}})) \sim \mathcal{N}(-1.3\,,0.2)$, which corresponds to being centred around 50~Myr. Choosing this regime to fit allows more prior weight to be placed on shorter prompt times as per the majority of the literature.

@@

The majority of works in the field have constrained the DTD power-law slope, and many also estimate its normalisation. Conversely, few have attempted to constrain the time after a burst of star-formation at which SNe Ia begin to explode -- the prompt time $t_{\mathrm{p}}$. In most cases, $t_{\mathrm{p}}$ has been fixed at some fiducial value such as 40~Myr, which is derived from the lifetime of 8~M$_{\odot}$ stars. \citet{Castrillo2020} included $t_{\mathrm{p}}$ (which they denote $\Delta$) in their fit, and find a value of $50_{-35}^{+100}$~Myr which is consistent with both 40~Myr as well as our best fitting value of $90 \pm 50$~Myr. \citet{Heringer2019} also present fits with varying prompt times, although they don't fit for the parameter itself. They find that varying $t_{\mathrm{p}}$ changes the recovered normalisation, which is consistent with the degeneracy seen in our results (Fig. \ref{fig:corner_beta_norm_tp}). We therefore report results that are consistent with the fiducial value of 40~Myr.

@@

Low mass galaxies are highly sensitive to environmental effects and may often be found in the so-called starburst state with abnormally high (above the SFMS) levels of star formation that could drive environmental evolution \citep[e.g.][]{Dekel1986,Marlowe1995,Tremonti2004,Lee2006}, although studies have shown that only 6 percent of dwarf galaxies are starbursts \citep{Lee2009}. It is possible that the averaging effects of the mass assembly model overestimate the true, instantaneous SFR for a significant fraction of the dwarf galaxies. Since the SN Ia rate is dominated by short delay times in these populations, an overestimated SFR in the model would lead to an overestimate in the SN Ia rate.

@@

One possible cause of the discrepant SN Ia rate at high stellar mass is the effect of stellar metallicity. Metallicity has been previously invoked to explain irregularities or divergences from the fiducial DTD \citep[e.g.][]{Strolger2010,Meng2011,Kistler2013}. Metallicity may affect the observed rate of SNe in two ways: firstly, it can affect the time taken for a star to evolve along the main sequence ($t_{\mathrm{MS}}$) -- Stellar metallicity has varying effects on the MS lifetime of stars \citep[e.g.][]{Georgy2013,Amard2020}, which is also dependent upon initial rotation and degree of mixing; secondly it can affect the time taken from WD formation to SN Ia explosion -- low metallicity stars should produce higher-mass WDs \citep[e.g.][]{Umeda1999,Marigo2007}, resulting in a higher SN Ia rate \citep{Kistler2013}.
